---
title: "diffSpliSE analysis"
author: "Craig I Dent"
date: "23/01/2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=6)
options(scipen=6)
```

Load Libraries
```{r}
library(edgeR)
library(dplyr)
library(tidyr)
library(ggplot2)
```

Read in the diffSpliSE output file
```{r}
#ENTER ABSOLUTE PATH TO diffSPLISE OUTPUT FILE HERE
#         V
inFILE <-""
#ENTER 'TARGET' FILE HERE
#             V
targetFILE <-""

#Read in data
SpliSER_out<-read.table(inFILE,row.names=NULL,header=TRUE,sep="\t")
SpliSER_out
targets <-read.delim(targetFILE, stringsAsFactors=FALSE)
Group <- factor(targets$Group)
Group2 <- rep(Group, each =2)
lib_size2 <- rep(targets$Library_size, each=2)
targets
```

Filter NA containing Rows (not including the Gene Coloumn)
```{r}
SpliSER_narm <- SpliSER_out[complete.cases(subset(SpliSER_out, select=-c(Gene))), ]
SpliSER_narm
```

Filter all those sites that dont change across all samples, irregardless of condition.
```{r}
SpliSER_SSE <- SpliSER_narm %>% dplyr:: select(grep("_SSE", names(SpliSER_narm)))
SpliSER_SSE$meanSSE <- rowMeans(SpliSER_SSE)
SpliSER_SSE

SpliSER_filt <- SpliSER_narm[SpliSER_SSE$meanSSE>0.05,] 
SpliSER_filt <- SpliSER_narm[SpliSER_SSE$meanSSE<0.95,]
SpliSER_filt

```

Make a DGEList dataframe
```{r}
counts <- data.matrix(SpliSER_filt %>% dplyr:: select(grep("_alpha|_beta", names(SpliSER_filt))), rownames.force = FALSE)
labels <- SpliSER_filt[1:4] 
counts
labels
y <- DGEList(counts= counts, genes= labels, lib.size = lib_size2, group=Group2)
y
```

Record Sample Groupings
```{r}
Eff <- rep(c("alpha","beta"), length(Group))
Eff <- factor(Eff, levels=c("alpha","beta"))
Sample2 <- rep(targets$Sample, each=2)
Sample2 <- factor(Sample2)
```
Fit
```{r}
Combo <- factor(paste(Eff,Group2,sep="."))
design <- model.matrix(~0+Combo)
colnames(design) <- levels(Combo)
design
y <- estimateDisp(y, design=design, trend="none")
y$common.dispersion
summary(y$prior.df)
fit<- glmFit(y,design)
```

Pull contrast out of design  matrix
```{r}
designCols <- as.character(colnames(design))
designCols

contrastList <- c("(",designCols[1],"-",designCols[3],")-(",designCols[2],"-",designCols[4],")")
contrastString <- paste(contrastList, sep = '')

myargs <- list(contr = contrastString, levels=design)
#contr=(alpha.Col-beta.Col)-(alpha.SC35.scl-beta.SC35.scl)
contrast <- do.call(makeContrasts,myargs)
contrast
lrt <- glmLRT(fit, contrast=contrast)
```
Record values for all splice-sites, and those with a significant p-value
```{r}
all<- topTags(lrt, n = 1000000, adjust.method = "BH",  p.value = 1.1) 
all

hits<- topTags(lrt, n = 1000000, adjust.method = "BH",  p.value = 0.05) 
hits

```

Merge statistical test data with rest of data
```{r}
hits_intersect <- merge(hits, SpliSER_filt, by=c("Region","Site","Gene")) 
drops<- c("logFC","logCPM","LR", "meanSSE")
hits_intersect <- hits_intersect[ , !(names(hits_intersect) %in% drops)]

#CURRENTLY THIS SECTION NEEDS TO BE UPDATED WITH YOUR NAMES (eg $Sample1Rep1_SSE)
hits_intersect$avg1 = (hits_intersect$ColA_SSE + hits_intersect$ColB_SSE +hits_intersect$ColC_SSE)/3
hits_intersect$avg2 = (hits_intersect$sc35.sclA_SSE + hits_intersect$sc35.sclB_SSE +hits_intersect$sc35.sclC_SSE)/3
hits_intersect$dSSE = hits_intersect$avg2 - hits_intersect$avg1
hits_intersect
```

Filter based on the average change in SSE between groups
```{r}
hits_SSEfiltered <- hits_intersect[abs(hits_intersect$dSSE) >= 0.1,]
hits_SSEfiltered[order(hits_SSEfiltered$FDR),]

```

Output Data
```{r}
#ENTER ABSOLUTE PATH TO OUTPUT FOR FULL-DATASET HERE
#                                   V
write.table(hits_intersect, file = "All.tsv", sep='\t', row.names = FALSE, col.names = TRUE, quote=FALSE)
#ENTER ABSOLUTE PATH TO OUTPUT FOR DIFF. SPLICED SITES HERE
#                                     V
write.table(hits_SSEfiltered, file = "BHhits_SSEFiltered.tsv", sep='\t', row.names = FALSE, col.names = TRUE, quote=FALSE)
```